\bf The Left Arm Bias in ITF Taekwon-Do Patterns \\ Chon-Ji to Juche

### 2010-12-29 article.tex
Fixed wording on movement in Ge-Baek
Added comment regarding the conclusion
Is it possible to say something useful about *why* there are more
blocks with the left than with the right arm given the data set?
Added details to the approach, reworded the data section, plus minor fixes
    * Removed references to Wikipedia
    * Fixed quote reference
    * Fixed relevant examples
    * Removed tables with planned data regarding height
    * Added source to the Venn diagrams

  Combined in a pattern, it allows the student to systematically deal with one
  or several imaginary opponents under various assumptions, using every
  available attacking and blocking tool from different directions.
  Furthermore, the techniques should also be evenly distributed between the
  left and the right side\cite{cyclo:vol1}.

@@

  In this article, therefore, we analyze hand techniques in patterns Chon-Ji
  to Juche using tags and then report our findings of a left arm bias.

@@

  \emph{Tags}, on the other hand, are non-hierarchical descriptive terms,
  keywords or labels that is attached to an object for later retrieval
  \cite{golder:ct}\cite{huang:tt}\cite{shirky:ontology}\cite{wikipedia:tags}.

@@

\SubSection{Set theory}
  A \emph{set} is a collection of distinct objects. It is a plurality thought
  of as a unit\cite{hausdorff:sets}. \emph{Set theory} is the branch of
  mathematics that studies sets\cite{wikipedia:tags}.

@@

  For example, $\{1,5,8\}$ is a set with three elements: 1, 5 and 8.
  $\{5,8,1\}$ is the same set, since the order of the elements doesn't matter.
  $\{1,1,1\}$, however, is not a valid set because the element 1 occurs
  repeatedly. Elements in a set can contain strings as well, so $\{loons,
  terns, penguins\}$ is also a valid set.

@@

  The terminology dataset is organized by pattern names. It contains a list of
  all techniques in patterns Chon-Ji to Juche using terminology. It also
  includes motion type on movements that are not performed in normal motion.

  The English description to every movement was registered manually after
  reviewing, and making minor corrections to, the terminology.

@@

  Each pattern movement is tagged on the basis of their English description
  from the condensed encyclopedia. In addition, suitable tags were attached to
  the dataset while performing the pattern to add necessary detail.

  The first movement in Chon-Ji, for instance, is to ``[m]ove the left foot to
  B, forming a left walking stance toward B while executing a low block to B
  with the left forearm\cite{cyclo:vol1}.'' Removing the verbose parts of the
  description we're left with:

@@

  Converting the above to tags gives us:

@@

  Tagging movements in such a formalized manner, and registering it in a
  database, enables us to retrieve the data we're interested in using concepts
  from set theory.

@@

\begin{itemize}
  \item
    \emph{Yul-Gok Tul}
    \begin{itemize}
      \item
        {\bf Movement 24}, \emph{Gunnun so ap palkup bandae taerigi}, is
        tagged with $\{front\_strike\}$ due to the English description of
        striking with the right front elbow.
      \item
        {\bf Movement 27}, \emph{Gunnun so ap palkup bandae taerigi}, is
        tagged with $\{front\_strike\}$ due to the English description of
        striking with the left front elbow.
    \end{itemize}
  \item
    \emph{Joong-Gun Tul}
    \begin{itemize}
      \item
        {\bf Movement 8}, \emph{Gunnun so wipalkup taerigi}, is tagged
        $\{upper\_strike, \: upper\_elbow\}$ since the English description
        says that the movement is ``a right upper elbow strike.''
      \item
        {\bf Movement 10}, \emph{Gunnun so wipalkup taerigi}, is tagged
        $\{upper\_strike, \: upper\_elbow\}$ since the English description
        says that the movement is ``a left upper elbow strike.''
      \item
        {\bf Movement 11}, \emph{Gunnun so sang sewo jirugi}, is tagged as a
        left and right technique, with $\{left\_technique, \:
        right\_technique\}$.
      \item
        {\bf Movement 12}, \emph{Gunnun so sang dwijibo jirugi}, is tagged as
        a left and right technique, with $\{left\_technique, \:
        right\_technique\}$.
      \item
        {\bf Movement 27}, \emph{Nachuo so sonbadak noollo makgi}, is tagged
        as a $\{middle\_technique, \: low\_technique\}$ because the left hand
        is in the middle section and the right hand, forming the pressing
        block, is in the low section of the body. Furthermore, the movement
        has been tagged as a $\{right\_technique, \: left\_technique\}$ to
        reflect that both hands are being used.
      \item
        {\bf Movement 29}, \emph{Nachuo so sonbadak noollo makgi}, is tagged
        as a $\{middle\_technique, \: low\_technique\}$ because the right hand
        is in the middle section and the left hand, forming the pressing
        block, is in the low section of the body. Furthermore, the movement
        has been tagged as a $\{right\_technique, \: left\_technique\}$ to
        reflect that both hands are being used.
    \end{itemize}
  \item
    \emph{Po-Eun Tul}
    \begin{itemize}
      \item
        {\bf Movement 6}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the left fore fist,''
        and the second one is a ``side front block with the right inner
        forearm.'' When searching, this movement will only register as one
        block.
      \item
        {\bf Movement 7}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the right fore
        fist,'' and the second one is a ``side front block with the left inner
        forearm.'' When searching, this movement will only register as one
        block.
      \item
        {\bf Movement 24}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the right fore
        fist,'' and the second one is a ``side front block with the left inner
        forearm.'' When searching, this movement will only register as one
        block.
      \item
        {\bf Movement 25}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the left fore fist,''
        and the second one is a ``side front block with the right inner
        forearm.'' When searching, this movement will only register as one
        block.
    \end{itemize}

@@

  \item
    \emph{Ge-Baek Tul}
    \begin{itemize}
      \item
        The height of the kick in {\bf movement 23}, \emph{Twimyo yopcha
        jirugi}, is not registered.
    \end{itemize}
  \item
    \emph{Choong-Jang Tul}
    \begin{itemize}
      \item
        {\bf Movement 19}, \emph{Moorup najunde apcha busigi}, is tagged with
        $\{front\_snap\_kick, \: knee\}$ which has an effect on the total of
        low front snap kicks.
      \item
        {\bf Movement 41}, \emph{Gunnun so sang sonkal soopyong taerigi}, is
        tagged with $\{knife\_hand, \: twin\_knife\_hand\}$.
    \end{itemize}
\end{itemize}

@@

@@

\begin{table}
  \centering
  \begin{tabular}{l|c|c|c} \hline \hline
    & High & Middle & Low \\ \hline

    Blocks    & x   & y   & z \\
    Punches   & x   & y   & z \\
    Strikes   & x   & y   & z \\
    Thrusts   & x   & y   & z \\
    \hline

  \end{tabular}
  \caption{Total number of techniques by height.}
  %\label{tab:table1}
  \end{table}

\begin{table}
  \centering
  \begin{tabular}{l|c|c|c} \hline \hline
    & High & Middle & Low \\ \hline

    Blocks    & x   & y   & z \\
    Punches   & x   & y   & z \\
    Strikes   & x   & y   & z \\
    Thrusts   & x   & y   & z \\
    \hline

  \end{tabular}
  \caption{Total number of techniques by height and arm.}
  %\label{tab:table1}
  \end{table}

% percent more useful than numbers?

@@

    \bibitem{wikipedia:sets}
      \emph{Set theory}, July 19, 2010. \\
      \url{http://en.wikipedia.org/wiki/Set_theory}

    \bibitem{wikipedia:tags}
      \emph{Tag (metadata)}, July 19, 2010. \\
      \url{http://en.wikipedia.org/wiki/Tag}

>>>>

  Combined in a pattern, it allows the student to systematically deal with one
  or several imaginary opponents under various assumptions, using every
  available attacking and blocking tool from different directions. In
  addition, the techniques should also be evenly distributed between the left
  and the right side\cite{cyclo:vol1}.

@@

  In this article we analyze hand techniques in patterns Chon-Ji to Juche
  using tags, and then report our findings of a left arm bias.

@@

  \emph{Tags}, on the other hand, are non-hierarchical descriptive terms,
  keywords or labels that is attached to an object for later retrieval
  \cite{golder:ct}\cite{huang:tt}\cite{shirky:ontology}.

@@

\SubSection{Set theory}
  A \emph{set} is a collection of distinct objects. It is ``a plurality
  thought of as a unit''\cite{hausdorff:sets}. \emph{Set theory}, then, is the
  branch of mathematics that studies sets.

@@

  For example, $\{1,5,8\}$ is a set with three elements: 1, 5 and 8.
  $\{5,8,1\}$ is the same set, since the order of the elements doesn't matter.
  $\{1,1,1\}$, however, is not a valid set because the element 1 occurs
  repeatedly. Elements in a set can contain strings as well, so $\{blocks,
  punches, strikes, thrusts\}$ is also a valid set.

@@

  The dataset is organized by pattern names. It contains a list of all
  techniques in patterns Chon-Ji to Juche. For each movement, the following
  data has been registered:

  \begin{enumerate}
    \item Movement number
    \item English description
    \item Terminology
    \item Tags
  \end{enumerate}

@@

  The movement number, together with the English description of each movement,
  were registered manually from the condensed encyclopedia. After reviewing
  and making minor corrections to the terminology data, it too was added.

  Suitable tags were attached to every movement by first extracting keywords
  from their English description, and second adding more detailed tags while
  performing the pattern.

  The first movement in Chon-Ji, for instance, is to ``[m]ove the left foot to
  B, forming a left walking stance toward B while executing a low block to B
  with the left forearm\cite{cyclo:vol1}.'' Removing the verbose parts of the
  description we are left with:

@@

  Supplementing the above with tags containing more information gives us:

@@

  Tagging movements in such a formalized manner, and registering it in a
  database, enables us to retrieve the data we are interested in using
  concepts from set theory.

@@

\begin{itemize}
  \item
    \emph{Yul-Gok Tul}
    \begin{itemize}
      \item
        {\bf Movement 24}, \emph{Gunnun so ap palkup bandae taerigi}, is
        tagged with $\{front\_strike\}$ due to the English description of
        striking with the right front elbow.
      \item
        {\bf Movement 27}, \emph{Gunnun so ap palkup bandae taerigi}, is
        tagged with $\{front\_strike\}$ due to the English description of
        striking with the left front elbow.
    \end{itemize}
  \item
    \emph{Joong-Gun Tul}
    \begin{itemize}
      \item
        {\bf Movement 8}, \emph{Gunnun so wipalkup taerigi}, is tagged
        $\{upper\_strike, \: upper\_elbow\}$ since the English description
        says that the movement is ``a right upper elbow strike.''
      \item
        {\bf Movement 10}, \emph{Gunnun so wipalkup taerigi}, is tagged
        $\{upper\_strike, \: upper\_elbow\}$ since the English description
        says that the movement is ``a left upper elbow strike.''
      \item
        {\bf Movement 11}, \emph{Gunnun so sang sewo jirugi}, is tagged as a
        left and right technique, with $\{left\_technique, \:
        right\_technique\}$.
      \item
        {\bf Movement 12}, \emph{Gunnun so sang dwijibo jirugi}, is tagged as
        a left and right technique, with $\{left\_technique, \:
        right\_technique\}$.
      \item
        {\bf Movement 27}, \emph{Nachuo so sonbadak noollo makgi}, is tagged as
        a $\{middle\_technique, \: low\_technique\}$ because the left hand is
        in the middle section and the right hand, forming the pressing block,
        is in the low section of the body. The movement has, additionally, been
        tagged as a $\{right\_technique, \: left\_technique\}$ to reflect that
        both hands are being used.
      \item
        {\bf Movement 29}, \emph{Nachuo so sonbadak noollo makgi}, is tagged as
        a $\{middle\_technique, \: low\_technique\}$ because the right hand is
        in the middle section and the left hand, forming the pressing block, is
        in the low section of the body. The movement has, additionally, been
        tagged as a $\{right\_technique, \: left\_technique\}$ to reflect that
        both hands are being used.
    \end{itemize}
  \item
    \emph{Po-Eun Tul}
    \begin{itemize}
      \item
        {\bf Movement 6}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the left fore fist,''
        and the second one is a ``side front block with the right inner
        forearm.'' When searching, this movement will only register as one
        block.
      \item
        {\bf Movement 7}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the right fore
        fist,'' and the second one is a ``side front block with the left inner
        forearm.'' When searching, this movement will only register as one
        block.
      \item
        {\bf Movement 24}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the right fore
        fist,'' and the second one is a ``side front block with the left inner
        forearm.'' When searching, this movement will only register as one
        block.
      \item
        {\bf Movement 25}, \emph{Annun so ap joomuk noollo makgi}, consist of
        two blocks. The first is a ``pressing block with the left fore fist,''
        and the second one is a ``side front block with the right inner
        forearm.'' When searching, this movement will only register as one
        block.
    \end{itemize}

@@

  \item
    \emph{Ge-Baek Tul}
    \begin{itemize}
      \item
        {\bf Movement 23}, \emph{Twimyo yopcha jirugi}; the height of the kick
        is not registered.
    \end{itemize}
  \item
    \emph{Choong-Jang Tul}
    \begin{itemize}
      \item
        {\bf Movement 19}, \emph{Moorup najunde apcha busigi}, is tagged with
        $\{front\_snap\_kick, \: knee\}$ which has an effect on the total of
        low front snap kicks.
      \item
        {\bf Movement 41}, \emph{Gunnun so sang sonkal soopyong taerigi}, is
        tagged with $\{knife\_hand, \: twin\_knife\_hand\}$.
    \end{itemize}
\end{itemize}

@@

% Is it possible to say anything about why there are more blocks with the left
% hand, than with the right?

@@

@@
